### 2020-04-18 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.Rmd
Experimenting with equations in README.md

When was the first case reported? When did the country close international arrivals? When did initial measures of social distancing were implemented? The lockdown?

@@

Our dataset includes information about the daily number of cases of COVID-19 reported in Spain \footnote{cases positively tested. Exclude sintomatic cases diagnostic by doctor with-out Reacción en Cadena de la Polimersa (PCR) test} at the provincial level (NUTIII in eurostat terminology) from March 13th to April 11th (30 days and 50 spatial units). The Spanish National Government publishes periodic updates at the regional level (NUTII) and the information is also released at the provincial level as part of a collaborative project. This information is compiled from several sources, mainly the official web pages of the Spanish regional goverments. In addition, we consider two sets of explanatory variables. The first one, and the focus of this research, is set of two environmental variables, namely temperature and humidity. The second set provides some relevant controls for multivariate analysis, and refers to economic and demographic attributes of the province. Table \ref{tab:descriptive-statistics} shows the descriptive statistics and the provencance of the data.

@@

The spatial Seemingly Unrelated Regressions (SUR) methodology has been a popular method to estimate are used in the context of...

The Seemingly Unrelated Regression equations model (SUR from now on) is a multivariate econometric formulation used in different fields when the data set is organized as cross-sections for different time periods. The basis of this approach is well-known since the initial works of @zellner1962 and now is a popular methodology included in several Econometrics textbook (e.g. @greene2003). To our knowledge, @Anselin1988spatial was the first text that introduced the term spatial SUR in reference to a model made of ‘an equation for each time period, which is estimated for a cross section of spatial units’ (p. 141). From this milestone, a huge researches used this methodology in spatial framework. @rey1999,@fingleton2007multi, @lauridsen2010, @legallo2006, @lopez2017spatial are well-known examples in this line.

>>>>

@@

Our dataset includes information about the daily number of cases of COVID-19 reported in Spain at the provincial level (NUTIII in Eurostat terminology) for the period between March 13th and April 11th, inclusive. This gives a temporal coverage of 30 days and spatial coverage of 50 units. For our purposes, we consider positive cases reported, but excluding symptomatic cases diagnosed by a doctor without a Polymerase Chain Reaction (PCR) test. The Spanish National Government publishes periodic updates at the regional level (NUTII) and the information is also released at the provincial level as part of a collaborative project [**by whom?**]. This information is compiled from several sources, mainly the official web pages of the Spanish regional goverments. In addition, we consider two sets of explanatory variables. The first one, and the focus of this research, is set of two environmental variables, namely temperature and humidity, which are collected from official sources (i.g., AEMET, the state meteorology agency, and MAPA, the ministry of agriculture, fisheries, and food). The second set provides some relevant controls for multivariate analysis, and refers to economic and demographic attributes of the province (also collected from official sources, i.e., INE, the national statistics institute). Table \ref{tab:descriptive-statistics} shows the descriptive statistics and the provenance of the data.

@@

The Seemingly Unrelated Regression equations model (SUR from now on) is a multivariate econometric formulation used in different fields when the data set is organized as cross-sections for different time periods. The basis of this approach is well-known since the initial works of Zellner [-@Zellner1962efficient] and now is a popular methodology included in several Econometrics textbook [e.g., @Greene2003econometric]. To our knowledge, @Anselin1988spatial was the first text that introduced the term spatial SUR in reference to a model made of ‘an equation for each time period, which is estimated for a cross section of spatial units’ (p. 141). From this milestone, a huge researches used this methodology in spatial framework. Rey and Montouri [-@Rey1999us], Fingleton [-@Fingleton2007multi], Lauridsen et al. [-@Lauridsen2010spatiotemporal], Le Gallo and Dall’Erba [-@Legallo2006evaluating], and Lopez et al. [-@Lopez2017spatial] are well-known examples in this line.

### 2020-04-18 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.tex
Experimenting with equations in README.md

  \begin{abstract}
  Spreading with astonishing speed, the novel SARS-CoV2 has swept the
  globe, causing enormous stress to health systems and prompting social
  distance guidelines and mandates to arrest its progress. While there is
  encouraging evidence that early public health interventions have slowed
  the spread of the virus, this has come at a high cost as the global
  economy is brought to its knees. How and when to ease restrictions to
  movement hinges in part on the question whether SARS-CoV2 will display
  seasonality associated with variations in temperature and humidity. In
  this research, we address this question by means of a spatial analysis
  of the incidence of COVID-19 in the provinces in Spain. Use of a spatial
  Seemingly Unrelated Regressions (SUR) approach allows us to model the
  incidence of reported cases of the disease per 100,000 population, as a
  function of temperature and humidity, while controlling for GDP per
  capita, population density, percentage of older adults in the
  population, and presence of mass transit systems. An interesting aspect
  of the spatial SUR approach is that it models incidence as a contagion
  process. Our results indicate that incidence of the disease is lower at
  higher temperatures. The evidence with respect to humidity is more
  mixed, with coefficients for this variable that are significant in only
  some equations. Our control variables also yield interesting insights,
  as population density and percentage of older adults display negative
  associations with incidence of COVID-19, whereas the presence of mass
  transit systems in the province is associated with a greater incidence
  of the disease.\\
  \end{abstract}

@@

Some of what is thought about the possible seasonality of COVID-19 is
based on analogies to the patterns of other known respiratory viruses.
However, de Ángel Solá et al.~(2020) note that ``not all seasonal
respiratory viruses experience the same spatiotemporal patterns''
(section 4). This urges caution when extrapolating from known viruses,
and indeed, the evidence in this respect is inconclusive. At a global
scale, whereas de Ángel Solá et al.~(2020) see less risk in the Caribean
Basin, Coelho et al.~(Coelho et al., 2020) warn that at least during the
exponential phase, expansion of the virus is not driven by climate.
Similarly, whereas Araujo and Naimi (2020) argue that spread of
SARS-CoV2 will likely be constrained by climate, Harbert et al.~(2020)
remain unconvinced that spatial modelling can currently discriminate the
distribution of the disease on the basis of climate, at least in the
United States. To further complicate matters, much of the relevant work
has yet to be peer-reviewed and therefore is open to change (see for
example the challenge of Harbert et al. (2020) to Araujo and Naimi
(2020)). In the United States, the National Academy of Sciences,
Engineering, and Medicine was engaged by the Office of the Executive for
guidance on this matter (see National Academies of Sciences, Engineering
and Medicine, 2020). Their conclusion summarizes the situation well (see
p.~6): ``Some limited data support a potential waning of cases in warmer
and more humid seasons, yet none are without major
limitations\ldots{}Additional studies as the SARS-CoV-2 pandemic unfolds
could shed more light on the effects of climate on transmission.''

@@

Our dataset includes information about the daily number of cases of
COVID-19 reported in Spain
\footnote{cases positively tested. Exclude sintomatic cases diagnostic by doctor with-out Reacción en Cadena de la Polimersa (PCR) test}
at the provincial level (NUTIII in eurostat terminology) from March 13th
to April 11th (30 days and 50 spatial units). The Spanish National
Government publishes periodic updates at the regional level (NUTII) and
the information is also released at the provincial level as part of a
collaborative project. This information is compiled from several
sources, mainly the official web pages of the Spanish regional
goverments. In addition, we consider two sets of explanatory variables.
The first one, and the focus of this research, is set of two
environmental variables, namely temperature and humidity. The second set
provides some relevant controls for multivariate analysis, and refers to
economic and demographic attributes of the province. Table
\ref{tab:descriptive-statistics} shows the descriptive statistics and
the provencance of the data.

@@

The spatial Seemingly Unrelated Regressions (SUR) methodology has been a
popular method to estimate are used in the context of\ldots{}

The Seemingly Unrelated Regression equations model (SUR from now on) is
a multivariate econometric formulation used in different fields when the
data set is organized as cross-sections for different time periods. The
basis of this approach is well-known since the initial works of
({\textbf{???}}) and now is a popular methodology included in several
Econometrics textbook (e.g.~({\textbf{???}})). To our knowledge, Anselin
(1988) was the first text that introduced the term spatial SUR in
reference to a model made of `an equation for each time period, which is
estimated for a cross section of spatial units' (p.~141). From this
milestone, a huge researches used this methodology in spatial framework.
({\textbf{???}}),({\textbf{???}}), ({\textbf{???}}), ({\textbf{???}}),
({\textbf{???}}) are well-known examples in this line.

@@

\noindent where\ldots{}.

@@

@@

@@

@@

>>>>

  \begin{abstract}
  Spreading with astonishing speed, the novel SARS-CoV2 has swept the
  globe, causing enormous stress to health systems and prompting social
  distance guidelines and mandates to arrest its progress. While there is
  encouraging evidence that early public health interventions have slowed
  the spread of the virus, this has come at a high cost as the global
  economy is brought to its knees. How and when to ease restrictions to
  movement hinges in part on the question whether SARS-CoV2 will display
  seasonality associated with variations in temperature and humidity. In
  this research, we address this question by means of a spatial analysis
  of the incidence of COVID-19 in the provinces in Spain. Use of a spatial
  Seemingly Unrelated Regressions (SUR) approach allows us to model the
  incidence of reported cases of the disease per 100,000 population, as a
  function of temperature and humidity, while controlling for GDP per
  capita, population density, percentage of older adults in the
  population, and presence of mass transit systems. An interesting aspect
  of the spatial SUR approach is that it models incidence as a contagion
  process. Our results indicate that incidence of the disease is lower at
  higher temperatures. The evidence with respect to humidity is more
  mixed, with coefficients for this variable that are significant in only
  some equations. Our control variables also yield interesting insights,
  as population density and percentage of older adults display negative
  associations with incidence of COVID-19, whereas the presence of mass
  transit systems in the province is associated with a greater incidence
  of the disease.
  \end{abstract}

@@

Some of what is thought about the possible seasonality of COVID-19 is
based on analogies to the patterns of other known respiratory viruses.
However, de Ángel Solá et al.~(2020) note that ``not all seasonal
respiratory viruses experience the same spatiotemporal patterns''
(section 4). This urges caution when extrapolating from known viruses,
and indeed, the evidence in this respect is inconclusive. At a global
scale, whereas de Ángel Solá et al.~(2020) see less risk in the Caribean
Basin, Coelho et al.~(Coelho et al., 2020) warn that at least during the
exponential phase, expansion of the virus is not driven by climate.
Similarly, whereas Araujo and Naimi (2020) argue that spread of
SARS-CoV2 will likely be constrained by climate, Harbert et al.~(2020)
remain unconvinced that spatial modelling can currently discriminate the
distribution of the disease on the basis of climate, at least in the
United States. To further complicate matters, much of the relevant work
has yet to be peer-reviewed and therefore is open to change (see for
example the challenge of Harbert et al. (2020) to Araujo and Naimi
(2020)). In the United States, the National Academy of Sciences,
Engineering, and Medicine was engaged by the Office of the Executive for
guidance on this matter (see National Academies of Sciences, Engineering
and Medicine, 2020). Their conclusion summarizes the situation well (see
p.~6): ``Some limited data support a potential waning of cases in warmer
and more humid seasons, yet none are without major
limitations\ldots Additional studies as the SARS-CoV-2 pandemic unfolds
could shed more light on the effects of climate on transmission.''

@@

Our dataset includes information about the daily number of cases of
COVID-19 reported in Spain at the provincial level (NUTIII in Eurostat
terminology) for the period between March 13th and April 11th,
inclusive. This gives a temporal coverage of 30 days and spatial
coverage of 50 units. For our purposes, we consider positive cases
reported, but excluding symptomatic cases diagnosed by a doctor without
a Polymerase Chain Reaction (PCR) test. The Spanish National Government
publishes periodic updates at the regional level (NUTII) and the
information is also released at the provincial level as part of a
collaborative project {[}\textbf{by whom?}{]}. This information is
compiled from several sources, mainly the official web pages of the
Spanish regional goverments. In addition, we consider two sets of
explanatory variables. The first one, and the focus of this research, is
set of two environmental variables, namely temperature and humidity,
which are collected from official sources (i.g., AEMET, the state
meteorology agency, and MAPA, the ministry of agriculture, fisheries,
and food). The second set provides some relevant controls for
multivariate analysis, and refers to economic and demographic attributes
of the province (also collected from official sources, i.e., INE, the
national statistics institute). Table \ref{tab:descriptive-statistics}
shows the descriptive statistics and the provenance of the data.

@@

The Seemingly Unrelated Regression equations model (SUR from now on) is
a multivariate econometric formulation used in different fields when the
data set is organized as cross-sections for different time periods. The
basis of this approach is well-known since the initial works of Zellner
(1962) and now is a popular methodology included in several Econometrics
textbook (e.g., Greene, 2003). To our knowledge, Anselin (1988) was the
first text that introduced the term spatial SUR in reference to a model
made of `an equation for each time period, which is estimated for a
cross section of spatial units' (p.~141). From this milestone, a huge
researches used this methodology in spatial framework. Rey and Montouri
(1999), Fingleton ({\textbf{???}}), Lauridsen et al.~(2010), Le Gallo
and Dall'Erba (2006), and Lopez et al.~(2017) are well-known examples in
this line.

@@

\noindent where\ldots.

@@

\leavevmode\hypertarget{ref-Greene2003econometric}{}%
Greene, W.H., 2003. Econometric analysis. Pearson Education India.

@@

\leavevmode\hypertarget{ref-Lauridsen2010spatiotemporal}{}%
Lauridsen, J., Bech, M., López, F., Maté, M., 2010. A spatiotemporal
analysis of public pharmaceutical expenditure. The Annals of Regional
Science 44, 299--314.

\leavevmode\hypertarget{ref-Legallo2006evaluating}{}%
Le Gallo, J., Dall'Erba, S., 2006. Evaluating the temporal and spatial
heterogeneity of the european convergence process, 1980--1999. Journal
of Regional Science 46, 269--288.

\leavevmode\hypertarget{ref-Lopez2017spatial}{}%
López, F.A., Martínez-Ortiz, P.J., Cegarra-Navarro, J.-G., 2017. Spatial
spillovers in public expenditure on a municipal level in spain. The
Annals of Regional Science 58, 39--65.

@@

\leavevmode\hypertarget{ref-Rey1999us}{}%
Rey, S.J., Montouri, B.D., 1999. US regional income convergence: A
spatial econometric perspective. Regional studies 33, 143--156.

@@

\leavevmode\hypertarget{ref-Zellner1962efficient}{}%
Zellner, A., 1962. An efficient method of estimating seemingly unrelated
regressions and tests for aggregation bias. Journal of the American
statistical Association 57, 348--368.

### 2020-04-18 README.Rmd
Experimenting with equations in README.md

\begin{equation}
\label{eq:sur-sim}
\begin{bmatrix}
y_1 \\ y_2 \\ \vdots \\ y_T
\end{bmatrix}
=
\begin{bmatrix}
X_1 & 0 & \cdots & 0 \\ 0 & X_2 & \cdots & 0 \\ \vdots & \vdots & \ddots & \vdots \\ 0 & 0 & \cdots & X_T
\end{bmatrix}
\
\begin{bmatrix}
\beta_1 \\ \beta_1 \\ \vdots \\ \beta_T
\end{bmatrix}
+
\begin{bmatrix}
\epsilon_1 \\ \epsilon_2 \\ \vdots \\ \epsilon_T
\end{bmatrix}
\end{equation}

>>>>

<a href="https://www.codecogs.com/eqnedit.php?latex=\begin{equation}&space;\label{eq:sur-sim}&space;\begin{bmatrix}&space;y_1&space;\\&space;y_2&space;\\&space;\vdots&space;\\&space;y_T&space;\end{bmatrix}&space;=&space;\begin{bmatrix}&space;X_1&space;&&space;0&space;&&space;\cdots&space;&&space;0&space;\\&space;0&space;&&space;X_2&space;&&space;\cdots&space;&&space;0&space;\\&space;\vdots&space;&&space;\vdots&space;&&space;\ddots&space;&&space;\vdots&space;\\&space;0&space;&&space;0&space;&&space;\cdots&space;&&space;X_T&space;\end{bmatrix}&space;\&space;\begin{bmatrix}&space;\beta_1&space;\\&space;\beta_1&space;\\&space;\vdots&space;\\&space;\beta_T&space;\end{bmatrix}&space;&plus;&space;\begin{bmatrix}&space;\epsilon_1&space;\\&space;\epsilon_2&space;\\&space;\vdots&space;\\&space;\epsilon_T&space;\end{bmatrix}&space;\end{equation}" target="_blank"><img src="https://latex.codecogs.com/gif.latex?\begin{equation}&space;\label{eq:sur-sim}&space;\begin{bmatrix}&space;y_1&space;\\&space;y_2&space;\\&space;\vdots&space;\\&space;y_T&space;\end{bmatrix}&space;=&space;\begin{bmatrix}&space;X_1&space;&&space;0&space;&&space;\cdots&space;&&space;0&space;\\&space;0&space;&&space;X_2&space;&&space;\cdots&space;&&space;0&space;\\&space;\vdots&space;&&space;\vdots&space;&&space;\ddots&space;&&space;\vdots&space;\\&space;0&space;&&space;0&space;&&space;\cdots&space;&&space;X_T&space;\end{bmatrix}&space;\&space;\begin{bmatrix}&space;\beta_1&space;\\&space;\beta_1&space;\\&space;\vdots&space;\\&space;\beta_T&space;\end{bmatrix}&space;&plus;&space;\begin{bmatrix}&space;\epsilon_1&space;\\&space;\epsilon_2&space;\\&space;\vdots&space;\\&space;\epsilon_T&space;\end{bmatrix}&space;\end{equation}" title="\begin{equation} \label{eq:sur-sim} \begin{bmatrix} y_1 \\ y_2 \\ \vdots \\ y_T \end{bmatrix} = \begin{bmatrix} X_1 & 0 & \cdots & 0 \\ 0 & X_2 & \cdots & 0 \\ \vdots & \vdots & \ddots & \vdots \\ 0 & 0 & \cdots & X_T \end{bmatrix} \ \begin{bmatrix} \beta_1 \\ \beta_1 \\ \vdots \\ \beta_T \end{bmatrix} + \begin{bmatrix} \epsilon_1 \\ \epsilon_2 \\ \vdots \\ \epsilon_T \end{bmatrix} \end{equation}" /></a>
